---
title: "Regression analysis"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(corrplot)
library(ggplot2)
library(faraway)
library(kableExtra)
library(stringr)

knitr::opts_chunk$set(
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
set.seed(1)

rename <- dplyr::rename
```

<style type="text/css">

h1.title {
  text-align: center;
}

</style>

Since our goal is to explore what kinds of features might influence and best predict our outcome - ladder score. I am considering a multiple linear regression to see how variables relate to the outcome. Once important question that I am interested in is how much does GDP impacts the Ladder Score, and does anything interact with GDP. Here we mainly focus on our latest data - happiness report from 2021.

&nbsp;

```{r include=FALSE, message=FALSE}
regression_df = 
  read_csv("data/world-happiness-report-2021.csv") %>% 
  janitor::clean_names() 
head(regression_df)
```

## Outcome: Ladder score 
- Rename the column name
```{r}
regression_df$country_name <- as.factor(regression_df$country_name)

regression_df$regional_indicator<- as.factor(regression_df$regional_indicator)

regression_df <- regression_df %>%
  rename(Country = country_name, Region = regional_indicator, Ladder = ladder_score,
         SD.Ladder = standard_error_of_ladder_score, GDP = logged_gdp_per_capita, Social.support = social_support,
         Life.exp = healthy_life_expectancy, Freedom = freedom_to_make_life_choices,
         Corruption = perceptions_of_corruption, Ladder.Dystopia = ladder_score_in_dystopia,
         EXP.LOG.GPD = explained_by_log_gdp_per_capita, EXP.SS = explained_by_social_support,
         EXP.HLE = explained_by_healthy_life_expectancy, EXP.FREE = explained_by_freedom_to_make_life_choices,
         EXP.GEN = explained_by_generosity, EXP.CUR = explained_by_perceptions_of_corruption,
         DYS.RES = dystopia_residual)

hap <- regression_df[,-c(4,5,6,13:20)]
colnames(hap)

hap <- hap %>%
  dplyr::mutate(Ladder = row_number())
head(hap)
```

- check NA
```{r}
sum(is.na(regression_df)) # there is no na
```

We first rename the column name for further use and check if there is NA.
It turns out that no NA is in our dataset. What's worth mentioning is that we transform the ladder score according to their rank(largest score as rank 1) to have a better understanding when comparing them.

&nbsp;

## Correlations between each variables
Now, we check the correlations between each of the variables, and then we print the correlation matrix and plot the correlation graph.
```{r}
num.var <- select_if(hap, is.numeric)
M<-cor(num.var)
M
corrplot(M)
```

&nbsp;

## Model selection
- First, we split the data into a training and validation set and star our analysis
```{r include=FALSE, message=FALSE, warning=FALSE}
#Data Partitioning
set.seed(10)

#select sample size of training data (75% of data for train, 25% for validation)
sample_size <- floor(0.75 * nrow(hap))

# using sample() to select sample of specific size without replacement
# seq_len() generates a regular sequence - defined here as nrow (number of rows) of our data
# size is setting how large the sequence is (111)
train_ind <- sample(seq_len(nrow(hap)), size = sample_size)

#creating train data based on train_ind
train <- hap[train_ind,]
#test data is what is left over after train_ind is taken out
test <- hap[-train_ind,]
```

- Then, We choose to use North America and ANZ as a reference group for comparision
```{r}
str(hap$Region)
levels(hap$Region)
hap$Region<-relevel(hap$Region, ref="North America and ANZ")
```
&nbsp;

###  Full model
```{r}
m1 <- lm(Ladder ~ ., data = train[,-1])
print(m1)
```
&nbsp;

### Null model
Looking at the null model, which is just the intercept (mean) of the outcome variable.
```{r}
null <- lm(Ladder ~ 1, data = train[,-1])
print(null)
```
&nbsp;

### Use stepwise comparision procedure
#### select the model with the lowest AIC, which uses a model’s maximum likelihood estimation (log-likelihood) as a measure of fit.
```{r}
step(m1)

m1.step <- lm(formula = Ladder ~ Region + GDP + Social.support + Freedom + Corruption, data = train[, -1])
```

```{r include=FALSE, message=FALSE, warning=FALSE}
print(m1.step)
```

Here we get the best performing model measuring by AIC. The predictor includes Region, GDP, Social.support, Freedom and Corruption. 

&nbsp;

#### Plot the the residuals to see how the model performs
```{r}
plot(m1.step)
```

Overall the residuals seem normally distributed and most assumptions seem to be maintained. However, the Scale-Location plot does show some heteroskedasticity, where the variance increases a bit around the center.

&nbsp;

## Add interaction term
### Explore the interaction
Here are some plots to initiate our hypothesis
```{r}
ggplot(train[1:10,], aes(x = reorder(Country, Ladder), y = Ladder, fill = GDP)) + 
  geom_bar(stat = "identity")+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  labs(title = "Top Ten Happiest Countries") + 
  ylab("Ladder Score")+
  xlab("Coutries")

ggplot(train[], aes(Ladder,GDP)) + 
  geom_point()
ggplot(train[], aes(Ladder,Freedom)) + 
  geom_point()
ggplot(train[], aes(GDP,Freedom)) + 
  geom_point()

ggplot(data=train[1:30,], aes(x=GDP, y=Ladder))+
  geom_line(size=2, aes(color=Freedom))

sd(hap$Freedom)+mean(hap$Freedom)
sd(hap$Freedom)-mean(hap$Freedom)

high.Free <- which(hap[,"Freedom"] >=0.9)
low.Free <- which(hap[,"Freedom"] <=.7)

sd(hap$GDP)+mean(hap$GDP)
sd(hap$GDP)-mean(hap$GDP)


high.GDP <- which(hap[,"GDP"] >= 10.6)
low.GDP <- which(hap[,"GDP"] <= 8.3)

ggplot(hap[high.GDP,], aes(x=Freedom, y=Ladder))+
  geom_line(size=2, aes(color=GDP))

ggplot(hap[low.GDP,], aes(x=Freedom, y=Ladder))+
  geom_line(size=2, aes(color=GDP))
```

It seems like GDP and Freedom may have some interaction. 

&nbsp;

### Model the interaction
```{r}
#interactions
int.mod <- lm(formula = Ladder ~ Region + GDP + Social.support + Freedom + Corruption + GDP*Freedom, data = train[, -1])

summary(int.mod)
```

We found that the slope of Freedom changes for every one increase in GDP, and as GDP increases, slope of freedom decreases.
Looking at the p-value, the value of both GDP and Freedom indicate they're insignificant after we adding the interaction term, but the interaction effect is significant now. At the same time, the p-value for Corruption indicates it's insignificant.

&nbsp;

## Our final model
```{r}
final_model <- lm(formula = Ladder ~ Region + GDP + Social.support + Freedom + GDP*Freedom, data = train[, -1])

print(final_model)
```

&nbsp;

#### Plot the final model's residuals to see the distribution
```{r}
plot(final_model)
```

&nbsp;

## Reshow the correlation score 
- Freedom and GDP's correlation
```{r}
cor(hap$GDP, hap$Freedom)
```

&nbsp;

-  Ladder and GDP's correlation
```{r}
cor(hap$GDP, hap$Ladder)
```

&nbsp;

## Conclusion based on the analysis
The correlations between GDP and Ladder Score seem very strong, and the initial regression model would suggest that GDP is a powerful predictor of the Happiness Ladder Score, but after adapting interaction term and refit model, we found that the relationship between Freedom and GDP cannot be ignored: these two seems to influence each other and we cannot eliminate either of them.